Explainability and Adversarial Robustness for RNNs

### 2019-12-20 paper/paper.tex
Fix 'attack type' inconsistencies

\begin{figure}[h]
\includegraphics[width=\columnwidth]{../plots/adv_comparison/adv_comparison_17.pdf}
\caption{Attack success ratios for both datasets and per attack family for CIC-IDS-2017. Depicted are flow detection accuracies for adversarial flows and for unmodified attack flows.}
\label{fig:adv_per_family}
\end{figure}

\subsection{Evaluation}
\autoref{fig:adv_per_family} depicts the performance for the investigated algorithms for both datasets and for each attack family in CIC-IDS-2017.
We used \gls{cw} with a tradeoff of 1 and, to provide a fair comparison, for \gls{fgsm} and \gls{pgd} we used an average $L_\infty$ distance as observed for \gls{cw}.

@@

%Such a metric should be easy to compute, easy to interpret and robust with respect to outliers.
We consider an adversary successful if they cause at least 50\% of all samples to be misclassified. In this case, the \gls{ars} is the average of the distances of the adversarial samples to the original samples for the 50\% of all samples which have the smallest distances. If the adversary doesn't manage to manipulate enough samples, the \gls{ars} is $\infty$.

@@

A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 99.43\% vs.99.65\%). We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.

@@

\subsection{Explainability Plots}
Knowing which features to investigate, it is important to analyze which feature values lead to classification as attack.
In literature, the use of Partial Dependency Plots (\gls{pdp}) has been proposed~\cite{friedman_greedy_2001}. To inspect attack types in detail, in this research we use a \textit{conditional} form of the \gls{pdp}. If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space, $f(\boldsymbol X) \in [0,1]$ the neural network's prediction and $c$ the attack class, we define the conditional \gls{pdp} for feature $i$  as
% \begin{equation} \label{eq:pdp}
% \text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
% \end{equation}
% To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
\begin{equation} \label{eq:pdp_conditional}
\text{\gls{pdp}}_{c,i}(w) = \mathbb E_{\boldsymbol X | C} \Big(f \left( X_1,\ldots,X_{i-1},w,X_{i+1},\ldots X_n \right) | c\Big),
\end{equation}
empirically approximating the conditional distribution by the observed subset that belongs to a certain class.

@@

Due to the use of a 5-tuple flow key, port numbers and the protocol identifier are constant for all packets in one flow.
Hence, we can consider the \gls{rnn} a regular classifier and reuse 
%which allows us to use
\glspl{pdp}, which have been proposed for non-recurrent classification methods, by plotting the \gls{rnn}'s flow prediction outcome over one of these  features.
The results (omitted for brevity) show that for some attack classes the port numbers play an important role. When looking at the class of benign traffic %(non-attack traffic)
it becomes apparent that traffic destined to a high destination port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers.

>>>>

\begin{figure}[h]
\includegraphics[width=\columnwidth]{../plots/adv_comparison/adv_comparison_17.pdf}
\caption{Attack success ratios for both datasets and per attack type for CIC-IDS-2017. Depicted are flow detection accuracies for adversarial flows and for unmodified attack flows.}
\label{fig:adv_per_family}
\end{figure}

\subsection{Evaluation}
\autoref{fig:adv_per_family} depicts the performance for the investigated algorithms for both datasets and for each attack type in CIC-IDS-2017.
We used \gls{cw} with a tradeoff of 1 and, to provide a fair comparison, for \gls{fgsm} and \gls{pgd} we used an average $L_\infty$ distance as observed for \gls{cw}.

@@

%Such a metric should be easy to compute, easy to interpret and robust with respect to outliers.
We consider an adversary successful if he can cause at least 50\% of all samples to be misclassified. In this case, the \gls{ars} is the average of the distances of the adversarial samples to the original samples for the 50\% of all samples which have the smallest distances. If the adversary doesn't manage to manipulate enough samples, the \gls{ars} is $\infty$.

@@

A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 99.43\% vs. 99.65\%). We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.

@@

\subsection{Explainability Plots}
Knowing which features to investigate, it is important to analyze which feature values lead to classification as attack.
In literature, the use of Partial Dependency Plots (\gls{pdp}) has been proposed~\cite{friedman_greedy_2001}. To inspect attack types in detail, in this research we use a \textit{conditional} form of the \gls{pdp}. If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space, $f(\boldsymbol X) \in [0,1]$ the neural network's prediction and $c$ the attack type, we define the conditional \gls{pdp} for feature $i$  as
% \begin{equation} \label{eq:pdp}
% \text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
% \end{equation}
% To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
\begin{equation} \label{eq:pdp_conditional}
\text{\gls{pdp}}_{c,i}(w) = \mathbb E_{\boldsymbol X | C} \Big(f \left( X_1,\ldots,X_{i-1},w,X_{i+1},\ldots X_n \right) | c\Big),
\end{equation}
empirically approximating the conditional distribution by the observed subset that belongs to a certain attack type.

@@

Due to the use of a 5-tuple flow key, port numbers and the protocol identifier are constant for all packets in one flow.
Hence, we can consider the \gls{rnn} a regular classifier and reuse 
%which allows us to use
\glspl{pdp}, which have been proposed for non-recurrent classification methods, by plotting the \gls{rnn}'s flow prediction outcome over one of these  features.
The results (omitted for brevity) show that for some attack types the port numbers play an important role. When looking at the \gls{pdp} for benign traffic samples, %(non-attack traffic)
it becomes apparent that traffic destined to a high destination port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers.
